---
title: "Assignment 2"
author: "Ishwar Choudhary , Shreya Banerji"
output: html_document
runtime: shiny
---
```{r setup, include=FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
#Q1
#install.packages("plotly")
library(plotly)
library(MASS)
require(data.table)
library(dplyr)

#Q4
#install.packages("ggmap")
library(ggmap)
library(ggplot2)
library(shiny)

#Q5
#install.packages("RColorBrewer")
#install.packages("tidyr")
#install.packages("textdata")
#install.packages("wordcloud")
library(dplyr)
library(tidytext)
library(wordcloud)
library(wordcloud)
library(reshape2)

```

```{r}
data=fread("fitness_data.csv")
d=na.omit(data)
sub_IDs=sort(unique(d$subID))       #obtaining all the ids
#plotting the graph
graph <- plot_ly(d, x = ~d$`timestamp (in seconds)`,
    transforms = list( 
      list(type = 'filter',
      target = d$subID,
      operation = '=',
      value = sub_IDs[1]
    ))) %>% 
add_trace(y = ~d$`Body Temperature (Celsius)`, name = "Body Temperatue",mode='lines+markers')  %>%
add_trace(y = ~d$`heartrate during activity (bpm)`, name = "heartrate during activity (bpm)",mode='lines+markers', visible = F)  %>%
add_trace(y = ~d$`Acceleration (x axis) in m/s2`, name = "Acceleration (x axis) in m/s2",mode='lines+markers', visible = F)  %>%
add_trace(y = ~d$`Acceleration (y axis) in m/s2`, name = "Acceleration (y axis) in m/s2",mode='lines+markers', visible = F)  %>%
add_trace(y = ~d$`Acceleration (z axis) in m/s2`, name = "Acceleration (z axis) in m/s2",mode='lines+markers', visible = F) %>%
add_trace(y = ~d$`Gyroscope (x axis) in rad/s`, name = "Gyroscope (x axis) in rad/s",mode='lines+markers',visible = F) %>%
add_trace(y = ~d$`Gyroscope (y axis) in rad/s`, name = "Gyroscope (y axis) in rad/s",mode='lines+markers', visible = F) %>%
add_trace(y = ~d$`Gyroscope (z axis) in rad/s`, name = "Gyroscope (z axis) in rad/s", mode='lines+markers',visible = F) %>%
add_trace(y = ~d[,12], name = "Magnetometer (x axis) in μT",mode='lines+markers',mode='lines+markers', visible = F) %>%
add_trace(y = ~d[,13], name = "Magnetometer (y axis) in μT",mode='lines+markers',mode='lines+markers', visible = F) %>%
add_trace(y = ~d[,14], name = "Magnetometer (z axis) in μT",mode='lines+markers', mode='lines+markers',visible = F) %>%
layout(
  title = "Fitness data",

  xaxis = list(
  title="x(timestamp in seconds)",
  rangeselector = list(buttons =
  list(list(
            count = 3, 
            stepmode = "backward"),
            list(step = "all"))),
            rangeslider = list(type = "time")),
  yaxis = list(title = "y"),
  updatemenus = list(
    list(
      type='dropdown',
      active=0,
      x=1.3,y = 0.9,
       buttons= list(
       list(method="restyle",args=list("transforms[0].value",sub_IDs[1]),
        label=sub_IDs[1]),
        list(method="restyle",args=list("transforms[0].value",sub_IDs[2]),
        label=sub_IDs[2]),
        list(method="restyle",args=list("transforms[0].value",sub_IDs[3]),
        label=sub_IDs[3]),
        list(method="restyle",args=list("transforms[0].value",sub_IDs[4]),
        label=sub_IDs[4]),
        list(method="restyle",args=list("transforms[0].value",sub_IDs[5]),
        label=sub_IDs[5]),
        list(method="restyle",args=list("transforms[0].value",sub_IDs[6]),
        label=sub_IDs[6]),
        list(method="restyle",args=list("transforms[0].value",sub_IDs[7]),
        label=sub_IDs[7]),
        list(method="restyle",args=list("transforms[0].value",sub_IDs[8]),
        label=sub_IDs[8]),
        list(method="restyle",args=list("transforms[0].value",sub_IDs[9]),
        label=sub_IDs[9])
    )),
    list(
      x=1.5,y = 0.75,
      buttons = list(
        list(method = "restyle",
        args = list("visible", list(TRUE, FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE)),
             label = "Body Temperature (Celsius)"),
 
        list(method = "restyle",
        args = list("visible",list(FALSE, TRUE ,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE)),
        label = "heartrate during activity (bpm)" ),
        
         list(method = "restyle",
          args = list("visible",list(FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE)),
          label = "Acceleration (x axis) in m/s2"  ),
        
         list(method = "restyle",
        args = list("visible",list(FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE)),
        label = "Acceleration (y axis) in m/s2"),
        
         list(method = "restyle",
          args = list("visible",list(FALSE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE)),
          label = "Acceleration (z axis) in m/s2"),
        
        list(method = "restyle",
          args = list("visible",list(FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,FALSE)),
             label = "Gyroscope (x axis) in rad/s"),
        list(method = "restyle",
          args = list("visible",list(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE)),
             label = "Gyroscope (y axis) in rad/s"),
        list(method = "restyle",
          args = list("visible",list(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE)),
             label = "Gyroscope (z axis) in rad/s"),
    
        list(method = "restyle",
         args = list("visible",list(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE)),
             label = "Magnetometer (x axis) in μT"),
        list(method = "restyle",
          args = list("visible",list(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE)),
             label = "Magnetometer (y axis) in μT"),
        list(method = "restyle",
          args = list("visible",list(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE)),
             label = "Magnetometer (z axis) in μT")
      ))))

graph
```


Insights gleaned from the plot:-
1.For most of the people,body temperature increases and then decreases with time.
2. The variation in all the other parameters is random.



<h2> Question 2</h2>
What is undersampling and oversampling? Consider the dataset subject.csv. 
Is there a case of undersampling or oversampling? 
If so, mention a technique to remedy the problem. Justify your answer.

```{r}

#Source: https://zyxo.wordpress.com/2008/12/30/oversampling-or-undersampling/

subject_df  = read.csv("subject.csv") #reading the csv file

print(subject_df)

#Checking for undersampling or oversampling in the csv
#We find that the current data is highly imbalanced. 

subject_aggregate_df = aggregate(subject_df$Dominant.hand, by=list(Sex=subject_df$Sex), FUN=length)
barplot(subject_aggregate_df$x,names.arg = subject_aggregate_df$Sex, ylab = "Population")

#The remedy to the problem is either to oversample the Female population or under sample the males
#Oversampling the females seems to be a better approach

```
Answer 2 :
The remedy to the problem is either to oversample the Female population or under sample the males
Oversampling the females is a better approach.




<h2> Question 3</h2>

There are various techniques for sampling data. 
Suggest a sampling technique that you think is ideal for the data in fitness_data.csv, and justify your choice.

```{r}
fitness_df = read.csv("fitness_data.csv")
activities_df = read.csv("activities.csv")

#Through these datasets, we notice that there exists only one non-medical field recorded, which is activity ID.
#There are a total of 17 activities being performed, and hence Cluster sampling technique can be utilised to sample the subjects.
```
Answer 3:
Through these datasets, we notice that there exists only one non-medical field recorded, which is activity ID.
There are a total of 17 activities being performed, and hence Cluster sampling technique can be utilised to sample the subjects.

<h2> Question 4</h2>

In August 2018, Election Commission of India made Lok sabha 2014(Lok Sabha-2014
data.csv) data public so that analysts can use it for 2019 Lok Sabha election. 
Provide a suitable visualisation that accounts for the distribution of votes across the country.

```{r}
#INPUTTING GOOGLE API KEY #INTERNET CONNECTION REQUIRED TO RUN THIS
register_google(key = 'AIzaSyDwCTPxbxMRd-nYr9b5zNs2FW8jbGluJe0')
map <- get_map(location = 'India', zoom = 5)
#let's read in the csv
election_df = read.csv("Lok Sabha-2014 data.csv")
#print(election_df)
```


```{r}
#Q4 SHINY APP
#We are visualising the spread of voters for each party as it tells us how much prevelance or success a party has had in a certain regioon.
#The circles on map locate the region and the radius of each circle is the corresponding margin of votes by which they won.
#We are using a shiny party selector drop down menu. 

ui<-shinyUI(fluidPage(
  titlePanel("Election Results"),
  # Your input selection
  sidebarPanel(
    selectInput("party", "Choose you Input:", choices = unique(election_df$PARTY))
  ),
  # Show the selected plot
  mainPanel(
    plotOutput("whichplot")
  )
))

server<-shinyServer(function(input, output) {
  
  # Fill in the spot we created for a plot
  output$whichplot <- renderPlot({
    
  bjp = subset(election_df,election_df$PARTY ==input$party,)
  points <- ggmap(map) + geom_point(aes(x = longitude, y = latitude,size = MARGIN), data = bjp ,alpha = .4,color= "orange")
  points <- points + scale_size_area(name = "MARGIN WIN")
  points
  })
})
shinyApp(ui, server)
```


```{r}
#IMPORTANT!
#NOTE AS SHINY APPLICATIONS ARE NOT SUPPORTED IN STATIC HTML PAGES
#THE ABOVE APP HAS BEEN IMPLEMENTED BELOW BUT WITHOUT THE ABILITY TO CHOSE A PARTY
#PLOTTING BY PARTY
#YOU CAN PUNCH IN ANY PARTY NAME AND IT WILL SHOW A MAP DEPICTING THE WIN REGION AND MARGIN OF WIN IS DEPICTED BY THE RADIUS
bjp = subset(election_df,election_df$PARTY =="Bharatiya Janata Party",)
cong = subset(election_df,election_df$PARTY =="Indian National Congress",)
points <- ggmap(map) + geom_point(aes(x = longitude, y = latitude,size = MARGIN), data = bjp ,alpha = .4,color= "orange") 
points <- points + scale_size_area(name = "MARGIN WIN")
points


#WE NEED TO PLOT A BAR GRAPH DENOTING THE NUMBER OF CONSTITUENCIES WON BY EVERY PARTY
party_aggregate = aggregate(election_df$PARTY, by=list(PARTY=election_df$PARTY), FUN=length)
party_aggregate = party_aggregate[order(-(party_aggregate$x)),] #- helps us sort in descending order
#print(party_aggregate)
par(mar=c(12,6.5,2.5,0.5),mgp=c(5,1,0)) 
barplot(party_aggregate$x,names.arg = party_aggregate$PARTY,las=2,col="ORANGE",main="Election Results",ylab = "Constituencies")
#xlab  missing as the party name overlaps it.
```

<h2>Question-5 </h2>
Many good Bollywood movies were released in 2019, one of them being Kabir Singh. The file tweets.txt contains what people have tweeted about this movie. Provide suitable visualization that depicts the generals sentiment of the audience.

```{r}

library(tidyr)
library(wordcloud)
library(reshape2)
library(dplyr)
library(tidytext)

filePath <- "tweets.txt"
text <- readLines(filePath)
text <- c(text)

text_df <- tibble(line = 1:12596, text = text)

text_df<-text_df %>%
  unnest_tokens(word, text)

text_df%>%
  count(word, sort = TRUE)

nrc_joy <- get_sentiments("nrc") %>%
  filter(sentiment == "joy")

text_df %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)

#showing the cloud

par(mar=c(0,1.5,0.5,0.5),mgp=c(10,1,0)) #do not change par

cloud = text_df %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("red", "blue"),
                   max.words = 100)

cloud

```


```{r}

```


